Nathan Greenslit Data Analysis Project

### 2023-04-05 products/manuscript/Manuscript.qmd
work cited

---
title: "Nathan Greenslit Data Analysis Project"
subtitle: ""
author: Nathan Greenslit
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

@@

The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year ([d'Almeida, 1986](https://journals.ametsoc.org/view/journals/apme/25/7/1520-0450_1986_025_0903_amfsdt_2_0_co_2.xml)). Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig. 1). These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each. Dust aerosols can harbor a wide diversity of bacteria ([Kellogg et al., 2004](https://link.springer.com/content/pdf/10.1023/B:AERO.0000032947.88335.bb.pdf)), fungi ([Ramírez Camejo et al., 2022](https://link.springer.com/article/10.1007/s10453-022-09752-9)), virus-like particles ([Griffin et al., 2001](https://link.springer.com/article/10.1023/A:1011868218901)), and minerals ([Formenti et al., 2003](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2002JD002648)). These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron ([Graham & Duce, 1967](https://www.sciencedirect.com/science/article/pii/0004698182901986), [Mills et al., 2004](https://pubmed.ncbi.nlm.nih.gov/15152251/), [Savoie et al., 1989](https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/JD094iD04p05069?casa_token=PMRaet9QSYsAAAAA:nb8yBorxdKaEWwGfI5aNILJ0D4h6drNu2YY1y1y8ExbQ03z59EUTQTOHoxinzsxRtR4jzsfSA73_Hig)). This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups ([Lenes et al., 2008](https://www.sciencedirect.com/science/article/pii/S0278434308000575), [Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113), [Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).

@@

Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens. Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming. The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V. cholerae*), shellfish-induced gastroenteritis (*V. parahaemolyticus, V. vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V. vulnificus*), and wound and ear infections  (*V. alginolyticus*, *V. vulnificus*) ([Baker-Austin et al., 2018](https://www.nature.com/articles/s41572-018-0005-8)). Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V. tubiashii, V. coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V. coralliilyticus*) ([Richards et al., 2015](https://pubmed.ncbi.nlm.nih.gov/25344234/), [Rosenberg & Falkovitz, 2004](https://www.annualreviews.org/doi/10.1146/annurev.micro.58.030603.123610), [Ben-Haim et al., 2003](https://pubmed.ncbi.nlm.nih.gov/12656189/)), and mortality in important aquaculture species such as penaeid shrimp (*V. parahaemolyticus*) and sea breams (*V. harveyi)* ([Tran et al., 2013](https://www.int-res.com/abstracts/dao/v105/n1/p45-55/), [Haldar et al., 2010](https://www.sciencedirect.com/science/article/pii/S0944501309001177)), among others.

>>>>

---
title: "Nathan Greenslit Data Analysis Project"
subtitle: ""
author: Nathan Greenslit
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: MADA.bib
---

@@

The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeida1986]. Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig. 1). These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each. Dust aerosols can harbor a wide diversity of bacteria [@kellogg2004], fungi [@ramirez2022], virus-like particles ([Griffin et al., 2001](https://link.springer.com/article/10.1023/A:1011868218901)), and minerals ([Formenti et al., 2003](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2002JD002648)). These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron ([Graham & Duce, 1967](https://www.sciencedirect.com/science/article/pii/0004698182901986), [Mills et al., 2004](https://pubmed.ncbi.nlm.nih.gov/15152251/), [Savoie et al., 1989](https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/JD094iD04p05069?casa_token=PMRaet9QSYsAAAAA:nb8yBorxdKaEWwGfI5aNILJ0D4h6drNu2YY1y1y8ExbQ03z59EUTQTOHoxinzsxRtR4jzsfSA73_Hig)). This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups ([Lenes et al., 2008](https://www.sciencedirect.com/science/article/pii/S0278434308000575), [Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113), [Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).

@@

Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens. Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming. The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V. cholerae*), shellfish-induced gastroenteritis (*V. parahaemolyticus, V. vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V. vulnificus*), and wound and ear infections  (*V. alginolyticus*, *V. vulnificus*) [@baker2018]. Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V. tubiashii, V. coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V. coralliilyticus*) ([Richards et al., 2015](https://pubmed.ncbi.nlm.nih.gov/25344234/), [Rosenberg & Falkovitz, 2004](https://www.annualreviews.org/doi/10.1146/annurev.micro.58.030603.123610), [@haim2003]), and mortality in important aquaculture species such as penaeid shrimp (*V. parahaemolyticus*) and sea breams (*V. harveyi)* ([Tran et al., 2013](https://www.int-res.com/abstracts/dao/v105/n1/p45-55/), [Haldar et al., 2010](https://www.sciencedirect.com/science/article/pii/S0944501309001177)), among others.
